---
title: "Exercise 17"
author: "Caitlin Jeffrey"
date: "March 28, 2019"
output: html_document
---


```{r}
poppy <- "C:/Users/caitl/OneDrive/Documents/BaseR/data/exercise_dat/bearclawpoppy.csv"
poppy<- read.csv("bearclawpoppy.csv", header = T, na.strings = c("", "99", "?", "na", "NA")); poppy$presab <- ifelse(poppy$PLANT %in% c("poppydead", "poppyalive", "poppy"), 1, ifelse(poppy$PLANT == "absence", 0, "NA")); poppy$presab <- as.factor(poppy$presab); str(poppy); levels(poppy$presab)
```

```{r}
poppy$presab=ifelse(poppy$presab=="0","0", ifelse(poppy$presab=="1", "1", "2")); presab<-poppy[,43]; vars<-poppy[,10:12]; new<-cbind(presab, vars); str(new)
```

```{r}
new<-lapply(new, as.numeric); str(new)
```

```{r}
for(i in 2:4){
  mn<-aggregate(new, by=list(new$presab), mean, na.rm=T)
  sd<-aggregate(new, by=list(new$presab), sd, na.rm=T)
  lngth<-aggregate(new, by=list(new$presab), length)
}
save(mn,sd,lngth, file = "newstats")
```



created 3 empty data objects, one for each statistical function, with appropriately named rows (presab poppy levels) and variable levels from poppy

```{r}
mn = array(0, dim=c(3,3)); sd = array(0, dim=c(3,3)); lng = array(0, dim=c(3,3)); dimnames(mn)[[1]]<-list("poppy$presab==0", "poppy$presab==1", "is.na(poppy$presab)"); dimnames(mn)[[2]]<-list("poppy$elev", "poppy$slope", "poppy$aspect"); dimnames(sd)[[1]]<-list("poppy$presab==0", "poppy$presab==1", "is.na(poppy$presab)"); dimnames(sd)[[2]]<-list("poppy$elev", "poppy$slope", "poppy$aspect"); dimnames(lng)[[1]]<-list("poppy$presab==0", "poppy$presab==1", "is.na(poppy$presab)"); dimnames(lng)[[2]]<-list("poppy$elev", "poppy$slope", "poppy$aspect")
```


```{r}
install.packages("PARSE", dependencies = T, repos = 'http://cran.us.r-project.org'); library(PARSE)
```
```{r}
install.packages("evaluate", dependencies = T, repos = 'http://cran.us.r-project.org'); library(evaluate)
```



```{r}
for (i in poppy[10:12])){
  for (j in c("poppy$elev", "poppy$slope", "poppy$aspect")){
    mn[i,j] <- aggregate(poppy[, i], by = list(poppy$presab, poppy$elev), FUN = mean, na.rm = T)
```


```{r}
for (i in c("poppy$presab==0", "poppy$presab==1", "is.na(poppy$presab)")){
  for (j in c("poppy$elev", "poppy$slope", "poppy$aspect")){
    mn[i,j] <- mean(eval(parse(text="j"))[eval(parse(text="i"))], na.rm=TRUE)
    sd[i,j] <- sd(eval(parse(text="j"))[eval(parse(text="i"))], na.rm=TRUE)
    lng[i,j] <-length(eval(parse(text="j"))[eval(parse(text="i"))])
  }
}
```
```{r}
for (i in c("poppy$presab==0", "poppy$presab==1", "is.na(poppy$presab)")){
  for (j in c("poppy$elev", "poppy$slope", "poppy$aspect")){
    mn[i,j] <- mean(eval(parse(text=j))[eval(parse(text=i))], na.rm=TRUE)
    sd[i,j] <- sd(eval(parse(text=j))[eval(parse(text=i))], na.rm=TRUE)
    lng[i,j] <-length(eval(parse(text=j))[eval(parse(text=i))])
  }
}
```

```{r}
for (i in c("poppy$presab==0", "poppy$presab==1", "is.na(poppy$presab)")){
  for (j in c("poppy$ELEV", "poppy$SLOPE", "poppy$ASPECT")){
    means[i,j] <- mean(eval(parse(text=j))[eval(parse(text=i))], na.rm=TRUE)
    sd[i,j] = sd(eval(parse(text=j))[eval(parse(text=i))], na.rm=TRUE)
    n[i,j] = length(eval(parse(text=j))[eval(parse(text=i))])
  } }
```



set working directory to powerpoint folder where m1-m4.csv are located; make blank list for file name output (in form of "m1.csv," etc), make list "ms" for output (in form "m1," etc"); loop to make list of m".csv and then read in csv files; then look at list of working objects in R to see that m".csv files were read in
 
```{r}
setwd(path.powerpoint)
files = list(); ms = list()
for (i in 1:4){
  files[i] <- paste('m', i, '.csv', sep = '')
  ms[i] <- paste('m', i, sep = '')
  assign(ms[[i]], read.csv(files[[i]]))
}; ls()
```
 

 
 
 
 
 
 
 