---
title: "cleaning codes"
author: "Caitlin Jeffrey"
date: "June 9, 2020"
output: html_document
---

```{r}
library(plyr)
```


## read in culture data csv

```{r}
setwd(path.exercise)
OREI<-read.csv("OREIculture_results_csv.csv", na.strings= ".", header = T)
str(OREI)
```

##find a specific value and row

```{r}
which(OREI$path_2 =="coryne"); OREI[3657,]
```

##pull out counts for each level of path type 1-4, rename with common column type, merge into 1 long list

```{r}
path1<-count(OREI$path_1); 
path2<-count(OREI$path_2); 
path3<-count(OREI$path_3);
path4<-count(OREI$path_4);
names(path1)[1] <- "pathID"; names(path1)[2] <- "freq";
names(path2)[1] <- "pathID"; names(path2)[2] <- "freq";
names(path3)[1] <- "pathID"; names(path3)[2] <- "freq";
names(path4)[1] <- "pathID"; names(path4)[2] <- "freq";
pathnum<-rbind(path1, path2, path3, path4)
```

##write out CSV

```{r}
write.csv(pathnum, "pathnum.csv", row.names = F)
```

##pull out counts for each level of path type 1-4 ONLY YES IMI, rename with common column type, merge into 1 long list

```{r}
path1a<-count(strictIMI$path_1); 
path2a<-count(strictIMI$path_2); 
path3a<-count(strictIMI$path_3);
path4a<-count(strictIMI$path_4);
names(path1a)[1] <- "pathID"; names(path1a)[2] <- "freq";
names(path2a)[1] <- "pathID"; names(path2a)[2] <- "freq";
names(path3a)[1] <- "pathID"; names(path3a)[2] <- "freq";
names(path4a)[1] <- "pathID"; names(path4a)[2] <- "freq";
pathnuma<-rbind(path1a, path2a, path3a, path4a)
```

##write out CSV

```{r}
write.csv(pathnuma, "pathnuma.csv", row.names = F)
```

##combine cowID, quarter, visit for unique ID for each data point

```{r}
OREI1<-(within(OREI, uniqueID <- paste(OREI$cow_id, OREI$quarter, OREI$visit, sep='_')))
```

##write out CSV with unique ID
```{r}
write.csv(OREI1, "OREI_culture_with_newID.csv", row.names = F)
```
 
 #checking blind quarters
```{r}
blind<-filter(all_trim, (first_blind=="yes" | second_blind=="yes" | third_blind=="yes" | fourth_blind == "yes"))
write.csv(blind, "blind.csv", row.names = T)
blind<-filter(never_all_withblind, (first_blind=="yes" | second_blind=="yes" | third_blind=="yes" | fourth_blind == "yes"))
```
 
 