---
title: "*Exploration of 40 herd bulk tank milk data*"
author: "Caitlin Jeffrey"
date: "August 6, 2021"
output: html_document
---
```{r}
library(ggplot2)
library(MASS)
library(ggpubr)
library(geoR)
library(sgeostat)
```


##***Discrete distributions***

###**Poisson**

* Range: [0,infinity]
* Parameters: size = number of events, rate = λ
* Interpretation: Distribution of events that occur during a fixed time interval or sampling effort with a constant rate of independent events; resembles normal with large λ, or exponential with small λ (example: over a 1 month period, what's the probability a bird lays a particular number of eggs; assumes constant rate of egg laying and eggs laid are independent of each other )

###**Binomial**

* Range: [0, # of trials]
* Parameters: size= number of trials; p = probability of positive outcome
* Interpretation: Distribution of number of successful independent dichotomous trials, with constant p (coin tossing, certain probability of getting heads)

###**Negative Binomial**

* Range: [0, infinity]
* Parameters: size=number of successes; p = probability of success
* Interpretation: Distribution of number of failures in a series of independent Bernouli trials, each with p = probability of a success. Generates a discrete distribution that is more heterogeneous ("overdispersed") than Poisson; (how many coins we have to toss before GETTING 10 heads?)

##***Continuous distributions***

###**Uniform**

* Range: [min,max]
* Parameters: min = minimum boundary; max = maximum boundary
* Interpretation: Distribution of a value that is equally likely within a specified range (any possible value prob is constant between two limits; special case of beta distributions)

###**Normal**

* Range: [neg infinity, infinity] 
* Parameters: mean = central tendency; SD = standard deviation
* Interpretation: Symmetric bell-shaped curve with unbounded tails

###**Gamma**

* Range: [0, infinity] (doens't go into 0, but continuous)
* Parameters: shape, scale
* Interpretation: mean=shape x scale, variance=shape x scale^2; generates a variety of shapes (including normal and exponential) for positive continuous variables; (unlike normal, NO NEG)

###**Beta**

* Range: [0,1] (can be rescaled to any range by simple multiplication and addition)
* Paramters: shape1, shape2
* Interpretation: if shape1 and shape 2 are integers, interpret as a coin toss, with shape1 = # of successes + 1, shape2 = # of failures + 1. Gives distribution of value of p, estimated from data, which can range from exponential through uniform through normal (but all are bounded). Setting shape1 and shape2 <1 yields u-shaped distributions. (dist of PROB of getting a certain outcome)

```{r}
btm<-read.csv("BTM_data_combined_7_1_2021.csv", na.strings= ".", header = T)
#str(btm)
```


### **Maximum likelihood estimators of the parameters**
* MLE's - probability of the data, given the hypothesis (standard would be null hypothesis - how probable are data given null is true)
* vs. what is probability of data given the parameters - how well does data fit given parameters of distribution - goodness of fit test
* vs. what is probability of the parameters, given the data - for a given data set, what is probability that certain set of parameters would be there- want parameter combinations that make it most likely to have come up with the data; maximize this probability (parameters given this data)

##***SCC***

```{r}
summary(btm$SCC_1.1000)
scc_data<-btm$SCC_1.1000
z <-fitdistr(scc_data,"normal")
print(z) # mean and SD, the MLE's (best fit for scc data; most likely if scc data is normally distributed)

# plot the density function for the normal data with these parameters and annotate with original data points (see smooth density normal curve, and where our data lie on that curve- does model fit data in reasonable way)

x <- 0:300000 # range of x values across where you want distribution
p_density <- dnorm(x=x,
                   mean=z$estimate["mean"],
                   sd=z$estimate["sd"]) # use mean and sd from actual scc data
qplot(x,p_density,geom="line") + annotate(geom="point",x=scc_data,y=0.0000001,color="springgreen4") + labs(title="SCC data against a normal distribution density curve",
       x="BTM SCC of 21 herds",
       y="Probability density of normal dist.") # see data points (x values) fall right in middle of probability mass curve

ggdensity(btm$SCC_1.1000, 
          main = "Density plot of BTM SCC",
          xlab = "SCC of 21 herds")

#Probability plot: The data are plotted against a theoretical distribution in such a way that the points should form approximately a straight line. Departures from this straight line indicate departures from the specified distribution. The correlation coefficient associated with the linear fit to the data in the probability plot is a measure of the goodness of the fit

#The P-P plot plots the cumulative probability of a variable against the cumulative probability of a particular distribution (e.g., normal distribution). After data are ranked and sorted, the corresponding z-score is calculated. This is the expected value that the score should have in a normal distribution. The scores are then themselves converted to z-scores. The actual z-scores are plotted against the expected z-scores. If the data are normally distributed, the result would be a straight diagonal line

#Q-Q plot is very similar to the P-P plot except that it plots the quantiles (values that split a data set into equal portions) of the data set instead of every individual score in the data

ggqqplot(btm$SCC_1.1000)

```

##***Raw CFU***

```{r}
summary(btm$RAW_CFU_ml_1.100)
rawcfu_data<-btm$RAW_CFU_ml_1.100
z <-fitdistr(rawcfu_data,"normal")
print(z) # mean and SD, the MLE's


x <- 0:100000 # range of x values across where you want distribution
p_density <- dnorm(x=x,
                   mean=z$estimate["mean"],
                   sd=z$estimate["sd"]) # use mean and sd from actual data
qplot(x,p_density,geom="line") + annotate(geom="point",x=rawcfu_data,y=0.0000001,color="deeppink2") + labs(title="Raw CFU count data against a normal distribution density curve",
       x="BTM Raw CFU count of 21 herds",
       y="Probability density of normal dist.") # see data points (x values) fall right in middle of probability mass curve

ggdensity(btm$RAW_CFU_ml_1.100, 
          main = "Density plot of BTM Raw CFU count",
          xlab = "Raw CFU of 21 herds")

ggqqplot(btm$RAW_CFU_ml_1.100)
```

##***Past CFU***

```{r}
summary(btm$PAST_CFU_ml_1.10)
pastcfu_data<-btm$PAST_CFU_ml_1.10
z <-fitdistr(pastcfu_data,"normal")
print(z) # mean and SD, the MLE's


x <- 0:2000 # range of x values across where you want distribution
p_density <- dnorm(x=x,
                   mean=z$estimate["mean"],
                   sd=z$estimate["sd"]) # use mean and sd from actual data
qplot(x,p_density,geom="line") + annotate(geom="point",x=pastcfu_data,y=0.0000001,color="deeppink1") + labs(title="Past. CFU count data against a normal distribution density curve",
       x="BTM Past. CFU count of 21 herds",
       y="Probability density of normal dist.") # see data points (x values) fall right in middle of probability mass curve

ggdensity(btm$PAST_CFU_ml_1.10, 
          main = "Density plot of BTM Past. CFU count",
          xlab = "Past. CFU of 21 herds")

ggqqplot(btm$PAST_CFU_ml_1.10)
```

##***PI CFU***

```{r}
summary(btm$PI__CFU_ml__1.1000)
picfu_data<-btm$PI__CFU_ml__1.1000
z <-fitdistr(picfu_data,"normal")
print(z) # mean and SD, the MLE's


x <- 0:2000000 # range of x values across where you want distribution
p_density <- dnorm(x=x,
                   mean=z$estimate["mean"],
                   sd=z$estimate["sd"]) # use mean and sd from actual data
qplot(x,p_density,geom="line") + annotate(geom="point",x=picfu_data,y=0.0000001,color="deeppink3") + labs(title="PI CFU count data against a normal distribution density curve",
       x="BTM PI CFU count of 21 herds",
       y="Probability density of normal dist.") # see data points (x values) fall right in middle of probability mass curve

ggdensity(btm$PI__CFU_ml__1.1000, 
          main = "Density plot of BTM PI CFU count",
          xlab = "PI CFU of 21 herds")

ggqqplot(btm$PI__CFU_ml__1.1000)
```

##***Total coliforms***

```{r}
summary(btm$Total_Coli__CFU_ml__1.1)
tcoli_data<-btm$Total_Coli__CFU_ml__1.1
z <-fitdistr(tcoli_data,"normal")
print(z) # mean and SD, the MLE's


x <- 0:30 # range of x values across where you want distribution
p_density <- dnorm(x=x,
                   mean=z$estimate["mean"],
                   sd=z$estimate["sd"]) # use mean and sd from actual data
qplot(x,p_density,geom="line") + annotate(geom="point",x=tcoli_data,y=0.0000001,color="deeppink4") + labs(title="Total coliform count data against a normal distribution density curve",
       x="BTM Total coliform count of 21 herds",
       y="Probability density of normal dist.") # see data points (x values) fall right in middle of probability mass curve

ggdensity(btm$Total_Coli__CFU_ml__1.1, 
          main = "Density plot of BTM Total coliform count",
          xlab = "Total coliform count of 21 herds")

ggqqplot(btm$Total_Coli__CFU_ml__1.1)
```


##***E. coli***

```{r}
summary(btm$E._coli__CFU_ml__1.1)
ecoli_data<-btm$E._coli__CFU_ml__1.1
z <-fitdistr(ecoli_data,"normal")
print(z) # mean and SD, the MLE's


x <- 0:20 # range of x values across where you want distribution
p_density <- dnorm(x=x,
                   mean=z$estimate["mean"],
                   sd=z$estimate["sd"]) # use mean and sd from actual data
qplot(x,p_density,geom="line") + annotate(geom="point",x=ecoli_data,y=0.0000001,color="mediumorchid") + labs(title="E. coli count data against a normal distribution density curve",
       x="BTM E. coli count of 21 herds",
       y="Probability density of normal dist.") # see data points (x values) fall right in middle of probability mass curve

ggdensity(btm$Total_Coli__CFU_ml__1.1, 
          main = "Density plot of BTM E. coli count",
          xlab = "E. coli count of 21 herds")

ggqqplot(btm$Total_Coli__CFU_ml__1.1)
```


##***Staph***

```{r}
summary(btm$Staph__CFU_ml__1.10)
staph_data<-btm$Staph__CFU_ml__1.10
z <-fitdistr(staph_data,"normal")
print(z) # mean and SD, the MLE's


x <- 0:200 # range of x values across where you want distribution
p_density <- dnorm(x=x,
                   mean=z$estimate["mean"],
                   sd=z$estimate["sd"]) # use mean and sd from actual data
qplot(x,p_density,geom="line") + annotate(geom="point",x=staph_data,y=0.0000001,color="mediumorchid2") + labs(title="Staph. count data against a normal distribution density curve",
       x="BTM Staph. count of 21 herds",
       y="Probability density of normal dist.") # see data points (x values) fall right in middle of probability mass curve

ggdensity(btm$Staph__CFU_ml__1.10, 
          main = "Density plot of BTM Staph count (St. Albans)",
          xlab = "Staph count of 21 herds")

ggqqplot(btm$Staph__CFU_ml__1.10)
```

##***Percent butterfat***

```{r}
summary(btm$perc_BF)
bf_data<-btm$perc_BF
z <-fitdistr(bf_data,"normal")
print(z) # mean and SD, the MLE's


x <- 0:10 # range of x values across where you want distribution
p_density <- dnorm(x=x,
                   mean=z$estimate["mean"],
                   sd=z$estimate["sd"]) # use mean and sd from actual data
qplot(x,p_density,geom="line") + annotate(geom="point",x=bf_data,y=0.0000001,color="royalblue1") + labs(title="Percent butterfat data against a normal distribution density curve",
       x="BTM Percent butterfat of 21 herds",
       y="Probability density of normal dist.") # see data points (x values) fall right in middle of probability mass curve

ggdensity(btm$perc_BF, 
          main = "Density plot of BTM % Butterfat",
          xlab = "% Butterfat of 21 herds")

ggqqplot(btm$perc_BF)
```

##***Percent protein***

```{r}
summary(btm$perc_Protein)
pro_data<-btm$perc_Protein
z <-fitdistr(pro_data,"normal")
print(z) # mean and SD, the MLE's


x <- 0:7 # range of x values across where you want distribution
p_density <- dnorm(x=x,
                   mean=z$estimate["mean"],
                   sd=z$estimate["sd"]) # use mean and sd from actual data
qplot(x,p_density,geom="line") + annotate(geom="point",x=pro_data,y=0.0000001,color="royalblue2") + labs(title="Percent protein data against a normal distribution density curve",
       x="BTM Percent protein of 21 herds",
       y="Probability density of normal dist.") # see data points (x values) fall right in middle of probability mass curve

ggdensity(btm$perc_Protein, 
          main = "Density plot of BTM % Protein",
          xlab = "% Protein of 21 herds")

ggqqplot(btm$perc_Protein)
```

##***Percent lactose***

```{r}
summary(btm$perc_Lactose)
lac_data<-btm$perc_Lactose
z <-fitdistr(lac_data,"normal")
print(z) # mean and SD, the MLE's


x <- 4:6 # range of x values across where you want distribution
p_density <- dnorm(x=x,
                   mean=z$estimate["mean"],
                   sd=z$estimate["sd"]) # use mean and sd from actual data
qplot(x,p_density,geom="line") + annotate(geom="point",x=lac_data,y=0.00000000000001,color="royalblue3") + labs(title="Percent lactose data against a normal distribution density curve",
       x="BTM Percent lactose of 21 herds",
       y="Probability density of normal dist.") # see data points (x values) fall right in middle of probability mass curve

ggdensity(btm$perc_Lactose, 
          main = "Density plot of BTM % Lactose",
          xlab = "% Lactose of 21 herds")

ggqqplot(btm$perc_Lactose)
```


##***Percent other solids***

```{r}
summary(btm$perc_Other.Solids)
solids_data<-btm$perc_Other.Solids
z <-fitdistr(solids_data,"normal")
print(z) # mean and SD, the MLE's


x <- 4:8 # range of x values across where you want distribution
p_density <- dnorm(x=x,
                   mean=z$estimate["mean"],
                   sd=z$estimate["sd"]) # use mean and sd from actual data
qplot(x,p_density,geom="line") + annotate(geom="point",x=solids_data,y=0.00000000000001,color="royalblue4") + labs(title="Percent other solids data against a normal distribution density curve",
       x="BTM Percent other solids of 21 herds",
       y="Probability density of normal dist.") # see data points (x values) fall right in middle of probability mass curve

ggdensity(btm$perc_Other.Solids, 
          main = "Density plot of BTM % Other solids",
          xlab = "% Other solids of 21 herds")

ggqqplot(btm$perc_Other.Solids)
```

##***MUN***

```{r}
summary(btm$MUN)
mun_data<-btm$MUN
z <-fitdistr(mun_data,"normal")
print(z) # mean and SD, the MLE's


x <- 0:25 # range of x values across where you want distribution
p_density <- dnorm(x=x,
                   mean=z$estimate["mean"],
                   sd=z$estimate["sd"]) # use mean and sd from actual data
qplot(x,p_density,geom="line") + annotate(geom="point",x=mun_data,y=0.00000000000001,color="firebrick2") + labs(title="MUN data against a normal distribution density curve",
       x="BTM MUN of 21 herds",
       y="Probability density of normal dist.") # see data points (x values) fall right in middle of probability mass curve

ggdensity(btm$MUN, 
          main = "Density plot of BTM MUN",
          xlab = "MUN of 21 herds")

ggqqplot(btm$MUN)
```

##***Minnesota coliforms***

```{r}
summary(btm$Coliforms)
mcoli_data<-btm$Coliforms
z <-fitdistr(mcoli_data,"normal")
print(z) # mean and SD, the MLE's


x <- 0:10 # range of x values across where you want distribution
p_density <- dnorm(x=x,
                   mean=z$estimate["mean"],
                   sd=z$estimate["sd"]) # use mean and sd from actual data
qplot(x,p_density,geom="line") + annotate(geom="point",x=mcoli_data,y=0.00000000000001,color="firebrick") + labs(title="Coliforms (Minn.) data against a normal distribution density curve",
       x="BTM Coliforms (Minn) of 21 herds",
       y="Probability density of normal dist.") # see data points (x values) fall right in middle of probability mass curve

ggdensity(btm$Coliforms, 
          main = "Density plot of BTM Coliforms (Minn)",
          xlab = "Coliform counts of 21 herds")

ggqqplot(btm$Coliforms)
```

##***Minnesota non-ag Streps***

```{r}
summary(btm$Non_ag_Strep)
strep_data<-btm$Non_ag_Strep
z <-fitdistr(strep_data,"normal")
print(z) # mean and SD, the MLE's


x <- 0:2000 # range of x values across where you want distribution
p_density <- dnorm(x=x,
                   mean=z$estimate["mean"],
                   sd=z$estimate["sd"]) # use mean and sd from actual data
qplot(x,p_density,geom="line") + annotate(geom="point",x=strep_data,y=0.0000001,color="firebrick2") + labs(title="Non-ag Streps (Minn.) data against a normal distribution density curve",
       x="BTM non-ag Streps (Minn) of 21 herds",
       y="Probability density of normal dist.") # see data points (x values) fall right in middle of probability mass curve

ggdensity(btm$Non_ag_Strep, 
          main = "Density plot of BTM Non-ag Streps (Minn)",
          xlab = "Non-ag Strep counts of 21 herds")

ggqqplot(btm$Non_ag_Strep)
```

##***Minnesota Staph. aureus***

```{r}
summary(btm$Staph_aureus)
sa_data<-btm$Staph_aureus
z <-fitdistr(sa_data,"normal")
print(z) # mean and SD, the MLE's


x <- 0:500 # range of x values across where you want distribution
p_density <- dnorm(x=x,
                   mean=z$estimate["mean"],
                   sd=z$estimate["sd"]) # use mean and sd from actual data
qplot(x,p_density,geom="line") + annotate(geom="point",x=sa_data,y=0.0000001,color="firebrick3") + labs(title="S. aureus (Minn.) data against a normal distribution density curve",
       x="BTM S. aureus (Minn) of 21 herds",
       y="Probability density of normal dist.") # see data points (x values) fall right in middle of probability mass curve

ggdensity(btm$Staph_aureus, 
          main = "Density plot of BTM Staph. aureus (Minn)",
          xlab = "Staph. aureus counts of 21 herds")

ggqqplot(btm$Staph_aureus)
```

##***Minnesota Staph. species***

```{r}
summary(btm$Staph_sp)
sp_data<-btm$Staph_sp
z <-fitdistr(sp_data,"normal")
print(z) # mean and SD, the MLE's


x <- 0:800 # range of x values across where you want distribution
p_density <- dnorm(x=x,
                   mean=z$estimate["mean"],
                   sd=z$estimate["sd"]) # use mean and sd from actual data
qplot(x,p_density,geom="line") + annotate(geom="point",x=sp_data,y=0.0000001,color="firebrick4") + labs(title="Staph sp. (Minn.) data against a normal distribution density curve",
       x="BTM Staph. sp. (Minn) of 21 herds",
       y="Probability density of normal dist.") # see data points (x values) fall right in middle of probability mass curve

ggdensity(btm$Staph_sp, 
          main = "Density plot of BTM Staph. species (Minn)",
          xlab = "Staph.species counts of 21 herds")

ggqqplot(btm$Staph_sp)
```


```{r}
#Code from Gotelli's notes

data <- c(0,0,0,0,0,0,1,1,1,2) # think this might fit poisson
z <- fitdistr(data,"poisson") # wants data, and what kind of dist you going to fit data to
print(z) # get a value for a parameter lambda - MLE based on these data is 0.5 (also gives SE, in parentheses) - pois only needs a lambda
sim_data <- rpois(n=10,lambda=z$estimate["lambda"])
head(sim_data) # simulation based on best estimate drawing from real data

# three steps to this activity:

# 1. select a statistical distribution that is appropriate for these data (integers, continuous, range of values it could be), determined by KIND of data you have
# 2. use fitdistr() function to find maximum likelihood estimates of the distribution parameters
# 3. use those parameters to visualize the probability density function ("d" family) and/or simulate additional data to use for other things


# --------------------------
# Example using real data 

frog_data <- c(30.15, 26.3, 27.5, 22.9, 27.8, 26.2)
z <-fitdistr(frog_data,"normal")
print(z) # mean and SD, the MLE's (best fit for frog data; most likely if frog data are what we are drawing from AND frog data is normally distributed)

# plot the density function for the normal data with these parameters and annotate with original data points (see smooth density normal curve, and where our data lie on that curve- does model fit data in reasonable way)

x <- 0:50 # range of x values across where you want distribution
p_density <- dnorm(x=x,
                   mean=z$estimate["mean"],
                   sd=z$estimate["sd"]) # use mean and sd from actual frog data
qplot(x,p_density,geom="line") + annotate(geom="point",x=frog_data,y=0.001) # see data points (x values) fall right in middle of probability mass curve

# do the same for the gamma distribution
z <-fitdistr(frog_data,"gamma")
print(z)
p_density <- dgamma(x=x,
                   shape=z$estimate["shape"],
                   rate=z$estimate["rate"])
qplot(x,p_density,geom="line") + annotate(geom="point",x=frog_data,y=0.001, color="red") #bound at 0




# --------------------------
# Example using real data + weird outlier!

outlier <- 0.05
frog_data <- c(30.15, 26.3, 27.5, 22.9, 27.8, 26.2)
frog_data <- c(frog_data,outlier)
z <-fitdistr(frog_data,"normal")
print(z) # mean and SD, the MLE's

# plot the density function for the normal data with these parameters and annotate with original data points

x <- 0:50
p_density <- dnorm(x=x,
                   mean=z$estimate["mean"],
                   sd=z$estimate["sd"])
qplot(x,p_density,geom="line") + annotate(geom="point",x=frog_data,y=0.001)

# do the same for the gamma distribution
z <-fitdistr(frog_data,"gamma")
print(z)
p_density <- dgamma(x=x,
                    shape=z$estimate["shape"],
                    rate=z$estimate["rate"])
qplot(x,p_density,geom="line") + annotate(geom="point",x=frog_data,y=0.001, color="red") #bound at 0; need to decide which dist is most appropriate; look at small data set compared to different distributions
```
