---
title: "stat201 homework3"
author: "Caitlin Jeffrey"
date: "June 16, 2020"
output: html_document
---
##module9 quiz
#In 1987, is there a difference in average life expectancy between Africa and the Americas?

```{r}
afam<-gapminder %>%
  filter(year==1987 & (continent=="Americas" | continent=="Africa"))

qqnorm(afam$lifeExp)
t.test(afam$lifeExp~afam$continent)
wilcox.test(afam$lifeExp~afam$continent)
```
#In 1987, what was the median life expectancy in Africa
```{r}
af87<-gapminder %>%
  filter(year==1987 & continent=="Africa")
```

#What would you estimate the average gdpPercap to be in Europe, in 1972?
```{r}
eur72<-gapminder %>%
  filter(year==1972 & continent=="Europe")
t.test(eur72$gdpPercap)
```

#In the gapminder dataset there are 5 continent represented, which continent has the most countries
```{r}
# how many unique countries does the data contain, by continent?
gapminder %>%
  group_by(continent) %>%
  summarise(n_obs = n(), n_countries = n_distinct(country))
```  

#Consider the gapminder data, suppose you would like to summarise the average life expectancy in 2007, by continent.
```{r}
gap07<-filter(gapminder, year==2007)
summarise(group_by(gap07, continent), mean=mean(lifeExp))
```

#you would like to create a new data.frame called "gapsum" which contains the 3 variables: continent, country, and average life expectancy (which is a variable you created that contains the average life expectancy across years for each country). Each country should be represented only once in the resulting data.frame

```{r}
gapsum<-gapminder %>% 
  group_by(country) %>%  
  mutate(meanle=mean(lifeExp))%>%  
  filter(year==2007) %>%  
  select(country, continent, meanle)
```



###module9 practice
#In 1952, is there a difference in average life expectancy between Asia and Europe? 
```{r}
eurasia52<-gapminder %>%
  filter(year==1952 &(continent=="Asia"|continent=="Europe"))

qqnorm(eurasia52$lifeExp)
t.test(eurasia52$lifeExp~eurasia52$continent)
wilcox.test(eurasia52$lifeExp~eurasia52$continent)
```
#In 2007, is there a difference in GDP among continents? Make a nice graph too.
```{r}
gdp<-gapminder %>%
  filter(year==2007)
summary(aov(gdp$gdpPercap~gdp$continent))
boxplot(gdp$gdpPercap~gdp$continent)
```
#Is there a difference in life expectancy in Africa between 1952 and 2007? 
```{r}
afr<-gapminder %>%
  filter((year==1952 | year==2007) & (continent=="Africa"))

qqnorm(afr$lifeExp); qqline(afr$lifeExp)
t.test(afr$lifeExp~afr$year, paired=T)
wilcox.test(afr$lifeExp~afr$year, paired=T)
```
#What would you estimate the average gdpPercap to be in Africa, in 1972? 
```{r}
afr<-gapminder %>%
  filter(continent=="Africa" & year==1972)
 t.test(afr$gdpPercap)
```

#For each country, what is the average lifeExp across all years (calculate one value that is the average of all years)?
```{r}
q5<-gapminder %>%
  group_by(country) %>%
  mutate(avglife=mean(lifeExp)) 

summarise(group_by(gapminder, country), avglife=mean(lifeExp))
```

#Using the variable you created in Q5, is there a difference in average lifeExp among continents?
```{r}
summary(aov(q5$avglife~q5$country))
```



##module9 video and practice questions

#read in csv's for video lecture

```{r}
demo<- read.csv("C:/Users/caitl/OneDrive/Documents/BaseR/data/exercise_dat/pt_demo.csv", header=T)
```

```{r}
labs<- read.csv("C:/Users/caitl/OneDrive/Documents/BaseR/data/exercise_dat/pt_labs.csv", header=T)
```

#Create a new variable in the demo data called "bpcat", which categorizes systolic blood pressure as: low (<120), med (>=120 & <132), or high (>=132)
#Write a command to see the frequencies of each category of "bpcat"
#Answer the question: "Is there a difference in cbmi between those who have hypertension (htn) and those who don't?" two groups distinct from each other, numeric variable, independent t-test

```{r}
demo$bpcat[demo$lsystolic<120]<-"low"
demo$bpcat[demo$lsystolic>=120&132]<-"med"
demo$bpcat[demo$lsystolic>=132]<-"high"
table(demo$bpcat, useNA="ifany")
```


```{r}
demo$htens[demo$bpcat=="low" | demo$bpcat=="medium"]<-"no"
demo$htens[demo$bpcat=="high"]<-"yes"
table(demo$htens, useNA="ifany")
#test first to see if data normally distributed
qqnorm(demo$cbmi)
t.test(cbmi~htens, data=demo)
#non-parametric
wilcox.test(cbmi~htens, data=demo)
```
#print first 6 obs

```{r}
head(demo)
```

```{r}
#Print observations 10 - 25 for columns 4 - 6 to your console
print( demo[
  c(10:25),
  c(4:6) ] )
#or
demo[10:25, 4:6]

#or
colnames(demo); print(demo[10:25, c("smoking", "smokecat", "smoker")])
```

#clean smoking category so NA's correctly classified
```{r}
table(demo$smoking)
#coerce from factor to character
demo$smoking<-as.character(demo$smoking)
demo$smoking[demo$smoking == ""] <-NA
unique(demo$smoking) #for characters, maybe level only for factors?
```

#merging data sets
```{r}
pts<-merge(demo,labs)
View(pts)
```

#using dyplyr
```{r}
library("dplyr")
```

#functions of dplyr
```{r}
glimpse(pts) #nice way to peek at structure of data frame
males<-filter(pts, sex==1) #filter is for choosing rows
forties<-filter(pts, age>=40 & age <50)
threevars<-select(pts, ptid, age, sex) #select is for choosing specific columns; pts is db, 3 variables
```

#Is RBC level associated with diabetes (dm) in current smokers?
```{r}
currentsmokersRBC<-filter(pts, smoker==1 & lab=="RBC")
chisq.test(currentsmokersRBC$dm, currentsmokersRBC$level)
fisher.test(currentsmokersRBC$dm, currentsmokersRBC$level)
```

#Is CMBI different between those with low systolic blood pressure and those with high systolic blood pressure (bpcat)?
```{r}
lowhi<-filter(pts, bpcat=="low" | bpcat=="high")
t.test(lowhi$cbmi~lowhi$bpcat)
```

#Is a1c level different among levels of WBC?

```{r}
wbc<-filter(pts, lab=="WBC")
summary(aov(wbc$a1c~wbc$level))
```

#What is the average age for those with both diabetes (dm) and hypertension (htn)?

```{r}
dmht<-filter(pts, lab=="RBC" & dm==1 & htn==1) #pick just one record per person
t.test(dmht$age)
mean(dmht$age)
```
########################
#summarise()- summary stats on variables within a data set; verb(object, newvariable=function(variable)
#group_by()-

```{r}
summarise(pts, avg=mean(age))
summarise(pts, mean=mean(cbmi, na.rm=T)) #need if have missing values
summarise(group_by(pts, sex), mean=mean(age)) #group the patients dataframe by the variable sex,then based on these groups summarize the variable age by calculating a mean- avg for each grouping variable
```

#make dataframe with only those who have htn and select 4 variables
```{r}
htn<-filter(pts, htn==1)
htn<-select(htn, ptid, age, ldiastolic, lsystolic)
```

#calc avg cmbi by diabetes status
```{r}
summarise(group_by(pts, dm), avg=mean(cbmi, na.rm=T))
```

#pipelining or chaining
```{r}
pts%>% #first select dataframe patients, AND THEN do...
  group_by(dm) %>% #group by diabetes, AND THEN
  summarise(avg=mean(cbmi, na.rm=T))
```

#calc median a1c level by hypertension status, JUST for those with diabetes
```{r}
pts%>% #first select dataframe patients, AND THEN do...
  filter(dm==1) %>%
  group_by(htn) %>%
  summarise(med=median(a1c, na.rm=T), mean=mean(a1c, na.rm=T))
```

#calc avg sys bp by smoking cat for individuals older than 35
```{r}
pts %>%
  filter(age>=35) %>%
  group_by(smokecat) %>%
  summarise(avgsbp=mean(lsystolic, na.rm=T))
```


###################
#arrange and mutate
#arrange will sort data
```{r}
pts1<-pts %>%
  arrange(age) #save new data frame as sorted by age
```

```{r}
pts1<-pts %>%
  arrange(sex, age)
```

#mutate will create new variables in your dataframe that are functions of existing variables
#newdataframe <- dataframe %>%
  #mutate(newvar=function(var))
  
```{r}
pts2<-pts%>%
  mutate(maxlabval=max(level)) #not super helpful bc max of ALL obs in set
```
  
```{r}
pts2<-pts%>%
  group_by(ptid) %>%
  mutate(maxlabval=max(level)) #grouped by patient, so max for each patient; creates a master variable for patient that has 3 records each in data set
```

#gapminder
```{r}
library("gapminder")
data(gapminder)
head(gapminder)
```











