---
title: "stat201 homework6"
author: "Caitlin Jeffrey"
date: "June 23, 2020"
output:
  pdf_document: default
  html_document: default
---

For this week's practice you'll need to write your own function, called bs(). It should produce a 95% bootstrap confidence interval for a mean. The user will provide a vector and the number of repetitions for the simulation.  

Your function should: 

* Draw a sample the same size as the vector, selecting from the vector itself, with replacement.  
* Compute the mean of the vector, and store it in another vector. 
* Repeat the steps above n times.  
* Compute the interval as the 2.5% and 97.5% quantiles of the vector of means.
* Print the interval on the console so that it looks something like this: "95% bootstrap interval for the mean:    -0.309  to  0.073"

Test your new function by creating a vector of 5000 z scores, and applying your function with 100 reps


#bootstrap confidence intervals
```{r}
z<-rnorm(5)
```

```{r}
#take repeated samples from z score- make bootstrap samples
s1<-sample(z, replace=T) #taken one bootstrap sample from population
s2<-sample(z, replace=T)
s3<-sample(z, replace=T)
s4<-sample(z, replace=T)
s5<-sample(z, replace=T)
```

* Draw a sample the same size as the vector, selecting from the vector itself, with replacement.  
* Compute the mean of the vector, and store it in another vector. 
* Repeat the steps above n times.  
* Compute the interval as the 2.5% and 97.5% quantiles of the vector of means.
* Print the interval on the console so that it looks something like this: "95% bootstrap interval for the mean:    -0.309  to  0.073"
```{r}
k<-5000
n<-100
bs<-function(k){
  for (z in 1:n) {
    z<-rnorm(k) #make a sample distribution pop of 5000 numbers, save as object z
    meanvec<-rep(NaN, k) #make a blank vector 5000 num long
    for (i in 1:k) {
      s<-sample(z, replace=T) #make 5000 sample pops by pulling from object z randomly 5000 times
      meanvec[i]<-mean(s)#take mean of s, put it in k spot in vector
    }
  }
  lowerlim<-quantile(meanvec, 0.025) #lower and upper limit of conf int
  upperlim<-quantile(meanvec, 0.975)
  print(paste("the lower confidence limit is is =", lowerlim, "and the upper confidence interval is =", upperlim))
}
bs(5000)
```

```{r}
lowerlim<-quantile(meanvec, 0.025) #lower and upper limit of conf int
upperlim<-quantile(meanvec, 0.975)
print(paste("the lower confidence limit is is =", lowerlim, "and the upper confidence interval is =", upperlim))
```


```{r}
k<-5000
n<-100
```


for (i in 1:10) {
  n<-rnorm(60, 30, 4)
  m_vec[i]<-mean(n)
}

```{r}
#then make a vector
meanvec<-rep(NaN, 5) 
meanvec[1]<-mean(s1)#take mean of s1, put it in first spot in vector
meanvec[2]<-mean(s2)
meanvec[3]<-mean(s3)
meanvec[4]<-mean(s4)
meanvec[5]<-mean(s5)
meanvec #vector filled with sampling means- 5 avgs from 5 bootstrap samplings
```

```{r}
#make lower and upper limit of conf int based on bootstrap means
lowerlim<-round(quantile(meanvec, 0.025),3) #to 3 dig
upperlim<-round(quantile(meanvec, 0.975),3)
```


```{r}
n<-100
bs(5000)
```


#Before watching the video, begin by creating a new data frame (z) in R that consists of 3 numeric variables: 
z1 - 50 randomly generated z scores
z2 - 50 randomly generated z scores
z3 - 50 randomly generated z scores

```{r}
z1<-rnorm(50)
z2<-rnorm(50)
z3<-rnorm(50) #don't specify mean and SD, default is 0 and 1
z<-data.frame(z1, z2, z3)
```

```{r}
z_stack<-stack(z) #stacking- combines variable in long format
summary(aov(values~ind, data=z_stack))
```

```{r}
View(mtcars)
mtc<-as.data.frame(t(mtcars)) #t function= transpose
```

```{r}
z<-rnorm(400)
y<-rnorm(400, 75, 8)
u<-round(runif(400)) #round to whole numbers only- uniform distribution, with a min of 0 and a max of 1
g<-round(runif(400,1,3)) #400 obs min 1 max 3
x<-as.data.frame(cbind(z,y,u,g))
```


```{r}
summary(x) #summary applies to entire dataframe, mean doesn't
mean(x$z) #mean need to use a variable inside dataframe
```


#Writing a new function (name of new func is "new"), takes some argument "k" and adds 3 to it- "K" can be called anything

```{r}
new<-function(k) {
  k+3
}
```

```{r}
new(4)
```

#function with 2 arguments
```{r}
new2<-function(k,p){
  k+p
}
```


```{r}
new2(6,4)
```

```{r}
des1<-function(k){
  summary(k)
  mean(k)
  sd(k)
  hist(k)
}
```

```{r}
des1(x$z) #why only return histogram? only returns last one to console
```

```{r}
des2<-function(k){ 
  print(summary(k))
  print(mean(k))
  print(sd(k))
  print(hist(k))
}
#will actually see results of function with print
```

```{r}
des2(x$z) #will actually see results of function
```

```{r}
#dealing with missing values- make any value bigger than 2 NA
x2<-x
x2$z[x2$z>2]<-NA
```

```{r}
des3<-function(k){ 
  print(summary(k))
  print(mean(k, na.rm=T))
  print(sd(k, na.rm=T))
  print(hist(k))
}
#dealing with missing values
```

```{r}
des3(x2$z) #run with missing values
```


```{r}
#adding labels to results
des4<-function(k){ 
  print(summary(k))
  m<-print(mean(k, na.rm=T))
  l<-print(sd(k, na.rm=T))
  print(paste("the std is =", l, "the mean is =", m))
}
```

```{r}
des4(x2$z)
```

```{r}
#calc mean from randomly selected sample of size y with a mean of 30 and an sd of 4
mean_go<-function(y){
  mean(rnorm(y,30,4))
}
```

```{r}
mean_go(25) #the avg from a sample of size 25
```
#now create a sampling distribution 
```{r}
x1<-mean_go(60)
x2<-mean_go(60)
x3<-mean_go(60)
x4<-mean_go(60)
x5<-mean_go(60)
x6<-mean_go(60)
x7<-mean_go(60)
x8<-mean_go(60)
x9<-mean_go(60)
x10<-mean_go(60)
x<-c(x1,x2,x3,x4,x5,x6,x7,x8,x9,x10)
hist(x)
```


#intro to loops!
#object = i, what you want to do to that object, 

```{r}
for (i in 1:4) {
  print(i)
}

for (i in 1:4) {
  print(i) #prints i, then moves on to the next thing
  print(i^2) #i, i^2, then 4 times
}
```

#create an empty vector
```{r}
m_vec<-rep(NaN, 10) #has 10 places, a blank vector

for (i in 1:10) {
  n<-rnorm(60, 30, 4)
  m_vec[i]<-mean(n)
}
hist(m_vec)
```

#now with a poisson distribution

```{r}
p_vec<-rep(NaN, 100) #has 100 places, a blank vector

for (i in 1:100) {
  p<-rpois(60, 3.5)
  m_vec[i]<-mean(p)
}

#hist(p_vec)
```

#bootstrap confidence intervals
```{r}
z<-rnorm(5)
t.test(z) #we think avg z score is... based on 5 scores
```

```{r}
#take repeated samples from z score- make bootstrap samples
s1<-sample(z, replace=T) #taken one bootstrap sample from population
s2<-sample(z, replace=T)
s3<-sample(z, replace=T)
s4<-sample(z, replace=T)
s5<-sample(z, replace=T)
```

```{r}
#then make a vector
meanvec<-rep(NaN, 5) 
meanvec[1]<-mean(s1)#take mean of s1, put it in first spot in vector
meanvec[2]<-mean(s2)
meanvec[3]<-mean(s3)
meanvec[4]<-mean(s4)
meanvec[5]<-mean(s5)
meanvec #vector filled with sampling means- 5 avgs from 5 bootstrap samplings
```

```{r}
#make lower and upper limit of conf int based on bootstrap means
lowerlim<-round(quantile(meanvec, 0.025),3) #to 3 dig
upperlim<-round(quantile(meanvec, 0.975),3)
```








