---
title: "subsetting helath categories"
author: "Caitlin Jeffrey"
date: "August 10, 2020"
output: html_document
---


##identifying cows that fall into health categories; "always infected"; "became infected"; "became infected and recovered"; "never infected"

```{r}
setwd(path.exercise)
#all_trim<-read.csv("all_trim.csv", na.strings= ".", header = T)
```


#always infected- need to ensure same path_ID, AND what about Stony Pond? go back and put them in- yes-yes
```{r}
#get cows from farms with only three visits (no CF) who are: yes-yes-yes
always_nocf<-filter(all_trim, (first_IMI_series=="yes" & second_IMI_series=="yes" & third_IMI_series=="yes") & (first_farm != "Choiniere" | second_farm != "Choiniere" | third_farm != "Choiniere" | fourth_farm !="Choiniere")) #162 observations

#get all CF cows, then always infected within CF
cf<-filter(all_trim, first_farm == "Choiniere" | second_farm == "Choiniere" | third_farm == "Choiniere" | fourth_farm == "Choiniere") #164 observations
cf_always<-filter(cf, (first_IMI_series=="yes" & second_IMI_series=="yes" & third_IMI_series=="yes" & fourth_IMI_series == "yes") | (first_IMI_series=="yes"  &  second_IMI_series=="yes" & third_IMI_series=="yes" & is.na(fourth_IMI_series)) | (is.na(first_IMI_series) & second_IMI_series=="yes" & third_IMI_series=="yes" & fourth_IMI_series=="yes")) #22 observations; how many visits with "YES" do we want to require? above coding catches 3 consecutive

#get all SP cows, then always infected within SP
sp<-filter(all_trim, first_farm == "StonyPond" | second_farm == "StonyPond")  #144 observations
sp_always<-filter(sp, (first_IMI_series=="yes" & second_IMI_series=="yes")) #19 observations

#combine CF and SP with rest of farms
always_nosp<-rbind(always_nocf, cf_always) #184 quarters
always_all<-rbind(always_nosp, sp_always) #203 observations

samepath_nosp<-filter(always_all, ((first_path_1==second_path_1) & (second_path_1==third_path_1)) | ((second_path_1==third_path_1) & (third_path_1==fourth_path_1))) #162 observations
samepath_sp<-filter(sp_always, (first_path_1==second_path_1)) #19 observations

samepath<-rbind(samepath_nosp, samepath_sp)
#not 100% perfect, could have become mixed infection or multiple pathogens isolated but not neccessarily called "path_1"; 181 observations
```

```{r}
always_csv<-table(samepath$second_farm) #79 tiestalls, 102 bedded pack; includes Stony Pond "yes"-"yes"
table(samepath$second_path_1) #144 CNS, 21 SA, 13 Strep, 2 Corynebac
write.csv(always_csv, "always_inf.csv", row.names = T)
not_same<-filter(chronic_all, ((first_path_1!=second_path_1) & (second_path_1!=third_path_1)) | ((second_path_1!=third_path_1) & (third_path_1!=fourth_path_1)))
```



#never infected- what about Stony Pond? Discluding non-negative quarters (called "no" in IMI_series vs. " ")
```{r}
#get cows from all farms with 3 visits who were no-no-no
never_nocf<-filter(all_trim, (first_IMI_series==""  & second_IMI_series==""  & third_IMI_series=="") & (first_farm != "Choiniere" | second_farm != "Choiniere" | third_farm != "Choiniere" | fourth_farm !="Choiniere")) #625 observations

#get object all with CF cows
cf<-filter(all_trim, first_farm == "Choiniere" | second_farm == "Choiniere" | third_farm == "Choiniere" | fourth_farm == "Choiniere") #164 observations

#get never infected CF cows
cf_never<-filter(cf, (first_IMI_series==""  &  second_IMI_series=="" & third_IMI_series=="" & fourth_IMI_series=="") | (first_IMI_series==""  &  second_IMI_series=="" & third_IMI_series=="" & is.na(fourth_IMI_series)) | (is.na(first_IMI_series) & second_IMI_series=="" & third_IMI_series=="" & fourth_IMI_series=="")) #60 observations

#get SP never infected
sp_never<-filter(sp, (first_IMI_series==""  &  second_IMI_series=="")) #90

#combine CF and SP with rest of farms
never_nosp<-rbind(never_nocf, cf_never) #685 observations
never_all_withblind<-rbind(never_nosp, sp_never) #775 observations

#remove any quarters that were or became blind
never_all_noblind<-filter(never_all_withblind, (first_blind=="" | is.na(first_blind)) & (second_blind=="" | is.na(second_blind)) & (third_blind=="" | is.na(third_blind)) & (fourth_blind=="" | is.na(fourth_blind))) #750 observations

# just checking using quarter codes
#qc_check<-filter(never_all_withblind, (first_quarter_code!="16" | is.na(first_quarter_code)) & (second_quarter_code!="16" | is.na(second_quarter_code)) & (third_quarter_code!="16" | is.na(third_quarter_code)) & (fourth_quarter_code!="16" | is.na(fourth_quarter_code))) #also 750 observations

never_3farms_qc<-filter(all_trim, (first_quarter_code=="3" | first_quarter_code=="4" | first_quarter_code=="5" | first_quarter_code=="6") & (second_quarter_code=="3" | second_quarter_code=="4" | second_quarter_code=="5" | second_quarter_code=="6") & (third_quarter_code=="3" | third_quarter_code=="4" | third_quarter_code=="5" | third_quarter_code=="6") & (first_farm != "Choiniere" | second_farm != "Choiniere" | third_farm != "Choiniere" | fourth_farm !="Choiniere")) #402 observations vs 625

cf_never_qc<-filter(cf, ((first_quarter_code=="3" | first_quarter_code=="4" | first_quarter_code=="5" | first_quarter_code=="6") & (second_quarter_code=="3" | second_quarter_code=="4" | second_quarter_code=="5" | second_quarter_code=="6") & (third_quarter_code=="3" | third_quarter_code=="4" | third_quarter_code=="5" | third_quarter_code=="6") & (fourth_quarter_code=="3" | fourth_quarter_code=="4" | fourth_quarter_code=="5" | fourth_quarter_code=="6")) | ((first_quarter_code=="3" | first_quarter_code=="4" | first_quarter_code=="5" | first_quarter_code=="6") & (second_quarter_code=="3" | second_quarter_code=="4" | second_quarter_code=="5" | second_quarter_code=="6") & (third_quarter_code=="3" | third_quarter_code=="4" | third_quarter_code=="5" | third_quarter_code=="6") & is.na(fourth_IMI_series)) | (is.na(first_IMI_series) & (second_quarter_code=="3" | second_quarter_code=="4" | second_quarter_code=="5" | second_quarter_code=="6") & (third_quarter_code=="3" | third_quarter_code=="4" | third_quarter_code=="5" | third_quarter_code=="6") & (fourth_quarter_code=="3" | fourth_quarter_code=="4" | fourth_quarter_code=="5" | fourth_quarter_code=="6")))

cf_never_qc1<-filter(cf, ((first_quarter_code=="3" | first_quarter_code=="4" | first_quarter_code=="5" | first_quarter_code=="6") & (second_quarter_code=="3" | second_quarter_code=="4" | second_quarter_code=="5" | second_quarter_code=="6") & (third_quarter_code=="3" | third_quarter_code=="4" | third_quarter_code=="5" | third_quarter_code=="6") & (fourth_quarter_code=="3" | fourth_quarter_code=="4" | fourth_quarter_code=="5" | fourth_quarter_code=="6"))) #11
                     
cf_never_qc2<-filter(cf, ((first_quarter_code=="3" | first_quarter_code=="4" | first_quarter_code=="5" | first_quarter_code=="6") & (second_quarter_code=="3" | second_quarter_code=="4" | second_quarter_code=="5" | second_quarter_code=="6") & (third_quarter_code=="3" | third_quarter_code=="4" | third_quarter_code=="5" | third_quarter_code=="6") & is.na(fourth_quarter_code))) #0
                     
cf_never_qc3<-filter(cf, (is.na(first_quarter_code) & (second_quarter_code=="3" | second_quarter_code=="4" | second_quarter_code=="5" | second_quarter_code=="6") & (third_quarter_code=="3" | third_quarter_code=="4" | third_quarter_code=="5" | third_quarter_code=="6") & (fourth_quarter_code=="3" | fourth_quarter_code=="4" | fourth_quarter_code=="5" | fourth_quarter_code=="6"))) #5

#16 observations
```

```{r}
never_csv<-table(never_all_noblind$second_farm) #404 tiestall, 371 bedded pack
write.csv(never_csv, "never_inf.csv", row.names = T)
```


#became infected
```{r}
became_inf<-filter(all_trim, (first_IMI_series==""  & second_IMI_series=="yes") | (second_IMI_series==""  & third_IMI_series=="yes") | (third_IMI_series==""  & fourth_IMI_series=="yes")) #147 observations

#compare to number observations using quarter codes, stricter definitions
became_inf_qc<-filter(all_trim, ((first_quarter_code=="3" | first_quarter_code=="4" | first_quarter_code=="5" | first_quarter_code=="6") & (second_IMI_series=="yes")) | ((second_quarter_code=="3" | second_quarter_code=="4" | second_quarter_code=="5" | second_quarter_code=="6") & (third_IMI_series=="yes")) | ((third_quarter_code=="3" | third_quarter_code=="4" | third_quarter_code=="5" | third_quarter_code=="6") & (fourth_IMI_series=="yes"))) #109 observations
```


```{r}
inf_csv<-table(became_inf$second_farm) #83 tiestall, 64 bedded pack
write.csv(inf_csv, "became_inf.csv", row.names = T)
table(became_inf$second_path_1)
table(became_inf$third_path_1)
table(became_inf$fourth_path_1)
```


#became infected, then recovered
```{r}
recovered<-filter(all_trim, (first_IMI_series==""  & second_IMI_series=="yes" & third_IMI_series=="") | (second_IMI_series==""  & third_IMI_series=="yes" & fourth_IMI_series=="")) #36 observations
table(recovered$second_path_1)

new_recovered<-filter(all_trim, (first_IMI_series=="yes"  & second_IMI_series=="") | (second_IMI_series=="yes"  & third_IMI_series=="") | (third_IMI_series=="yes" & fourth_IMI_series=="")) #130 observations, less strict- any quarter going from "yes" to "no"

yesnono_recovered<-filter(all_trim, (((first_IMI_series=="yes" & second_IMI_series=="" & third_IMI_series=="") | (second_IMI_series=="yes"  & third_IMI_series=="" & fourth_IMI_series=="")) & ((first_contam != "yes" | is.na(first_contam)) & second_contam != "yes" & third_contam != "yes") & (fourth_contam != "yes" | is.na(fourth_contam)))) #63 observations

#compare to number observations using quarter codes
recovered_qc<-filter(all_trim, ((first_quarter_code=="3" | first_quarter_code=="4" | first_quarter_code=="5" | first_quarter_code=="6")  & second_IMI_series=="yes" & (third_quarter_code=="3" | third_quarter_code=="4" | third_quarter_code=="5" | third_quarter_code=="6")) | ((second_quarter_code=="3" | second_quarter_code=="4" | second_quarter_code=="5" | second_quarter_code=="6")  & third_IMI_series=="yes" & (fourth_quarter_code=="3" | fourth_quarter_code=="4" | fourth_quarter_code=="5" | fourth_quarter_code=="6"))) #21 observations

new_recovered_qc<-filter(all_trim, (first_IMI_series=="yes"  & (second_quarter_code=="3" | second_quarter_code=="4" | second_quarter_code=="5" | second_quarter_code=="6")) | (second_IMI_series=="yes"  & (third_quarter_code=="3" | third_quarter_code=="4" | third_quarter_code=="5" | third_quarter_code=="6")) | (third_IMI_series=="yes" & (fourth_quarter_code=="3" | fourth_quarter_code=="4" | fourth_quarter_code=="5" | fourth_quarter_code=="6"))) #108 observations vs 130

yesnono_recovered_qc<-filter(all_trim, (first_IMI_series=="yes"  & (second_quarter_code=="3" | second_quarter_code=="4" | second_quarter_code=="5" | second_quarter_code=="6") & (third_quarter_code=="3" | third_quarter_code=="4" | third_quarter_code=="5" |third_quarter_code=="6")) | (second_IMI_series=="yes"  & (third_quarter_code=="3" | third_quarter_code=="4" | third_quarter_code=="5" | third_quarter_code=="6")) & (fourth_quarter_code=="3" | fourth_quarter_code=="4" | fourth_quarter_code=="5" | fourth_quarter_code=="6")) #54 observations vs 63
```

```{r}
old_recovered<-table(recovered$second_farm) #24 tiestall, 12 bedded packs
new_recovered<-table(new_recovered$second_farm) #65 tiestall, #65 bedded 
write.csv(old_recovered, "old_recovered.csv", row.names = T)
write.csv(new_recovered, "new_recovered.csv", row.names = T)
yyn_recovered<-table(yesnono_recovered$second_farm)
write.csv(yyn_recovered, "ynn_recovered.csv", row.names = T)
```


