---
title: "subsetting IMI"
author: "Caitlin Jeffrey"
date: "June 5, 2020"
output: html_document
---
```{r}
library(dplyr)
library(data.table)
```


## read in culture data csv

```{r}
setwd(path.exercise)
OREI<-read.csv("OREIculture_results_csv.csv", na.strings= ".", header = T)
str(OREI)
```


### reorganize data set so all obs for 1 cowquarter are lined up

```{r}
setwd(path.exercise)
OREI<-read.csv("OREIculture_results_csv.csv", na.strings= ".", header = T)
str(OREI)

#make SCC column numeric
OREI$scc<-as.numeric(as.character(OREI$scc))
```

#first, make 4 data sets- one for each visit
```{r}
library(dplyr); library(data.table)
setwd(path.exercise)
OREI<-read.csv("OREIculture_results_csv.csv", na.strings= ".", header = T)
#use dplyr and data.table functions, which seem to conflict some, need to read in OREI each time
first<-filter(OREI, visit=="first") #select only FIRST visit for each farm
oldnames<-c(colnames(first)) #make list of column names and save as object
newnames1<- paste("first_", oldnames, sep="")
setnames(first, old = c(oldnames), new = c(newnames1))
names(first)[names(first)=="first_cow_id"] <- "cow_id"
names(first)[names(first)=="first_quarter"] <- "quarter" #put back to original so can merge later
first$cow_id<-paste(first$cow_id, first$first_farm, sep='_') #so when go to merge, won't get screwy with same cow names at different farms!
first$first_uniqueID<-paste(first$cow_id, first$quarter, sep='_') #realized unique ID's in CSV aren't unique bc multiple animals same name different farm
```

```{r}
setwd(path.exercise)
OREI<-read.csv("OREIculture_results_csv.csv", na.strings= ".", header = T)
second<-filter(OREI, visit=="second") #select only SECOND visit for each farm
newnames2<- paste("second_", oldnames, sep="") #can still use oldnames, bc are same for all observations in OREI
setnames(second, old = c(oldnames), new = c(newnames2))
names(second)[names(second)=="second_cow_id"] <- "cow_id"
names(second)[names(second)=="second_quarter"] <- "quarter" #put back to original so can merge later
second$cow_id<-paste(second$cow_id, second$second_farm, sep='_')
second$second_uniqueID<-paste(second$cow_id, second$quarter, sep='_')
```


```{r}
setwd(path.exercise)
OREI<-read.csv("OREIculture_results_csv.csv", na.strings= ".", header = T)
third<-filter(OREI, visit=="third") #select only THIRD visit for each farm
newnames3<- paste("third_", oldnames, sep="") #can still use oldnames, bc are same for all observations in OREI
setnames(third, old = c(oldnames), new = c(newnames3))
names(third)[names(third)=="third_cow_id"] <- "cow_id"
names(third)[names(third)=="third_quarter"] <- "quarter" #put back to original so can merge later
third$cow_id<-paste(third$cow_id, third$third_farm, sep='_')
third$third_uniqueID<-paste(third$cow_id, third$quarter, sep='_')
```

```{r}
#setwd(path.exercise)
OREI<-read.csv("OREIculture_results_csv.csv", na.strings= ".", header = T)
fourth<-filter(OREI, visit=="fourth") #select only FOURTH visit for each farm
newnames4<- paste("fourth_", oldnames, sep="") #can still use oldnames, bc are same for all observations in OREI
setnames(fourth, old = c(oldnames), new = c(newnames4))
#names(fourth)[names(fourth)=="fourth_cow_id"] <- "cow_id"
#names(fourth)[names(fourth)=="fourth_quarter"] <- "quarter" #put back to original so can merge later
fourth$fourth_cow_id<-paste(fourth$fourth_cow_id, fourth$fourth_farm, sep='_')
fourth$fourth_uniqueID<-paste(fourth$fourth_cow_id, fourth$fourth_quarter, "fourth", sep='_')
#fourth<-select(fourth, -c(fourth_cow_id))
```



##merge all 4 visits so obs for 1 cow all in one row
```{r}
onetwo <- merge(first, second, by=c("cow_id","quarter"), all = TRUE) #should be NA for cows with missing observations
onetwothree <-merge(onetwo,third, by=c("cow_id", "quarter"), all = TRUE)
#onetwothree has all farms, visits 1-3
```

```{r}
onetwothree$fourth_uniqueID<-paste(onetwothree$cow_id, onetwothree$quarter, "fourth", sep='_') #populate fourth_unique_ID so can index when merge fourth visit in
all<-merge(onetwothree, fourth, by=c("fourth_uniqueID"), all = TRUE) #make a data frame with all 4 visits
```

```{r}
#two cows from choiniere were new 4th visit, so need to manually populate their cow id

all$cow_id[all$fourth_cow_id =="65_Choiniere"] = "65_Choiniere"
all$quarter[all$fourth_uniqueID=="65_Choiniere_LH_fourth"] = "LH"
all$quarter[all$fourth_uniqueID=="65_Choiniere_LF_fourth"] = "LF"
all$quarter[all$fourth_uniqueID=="65_Choiniere_RH_fourth"] = "RH"
all$quarter[all$fourth_uniqueID=="65_Choiniere_RF_fourth"] = "RF"
all$cow_id[all$fourth_cow_id =="87_Choiniere"] = "87_Choiniere"
all$quarter[all$fourth_uniqueID=="87_Choiniere_LH_fourth"] = "LH"
all$quarter[all$fourth_uniqueID=="87_Choiniere_LF_fourth"] = "LF"
all$quarter[all$fourth_uniqueID=="87_Choiniere_RH_fourth"] = "RH"
all$quarter[all$fourth_uniqueID=="87_Choiniere_RF_fourth"] = "RF"
```


```{r}
#can drop columns fourth cow id, fourth quarter
all_trim<-select(all, -c(fourth_cow_id, fourth_quarter, fourth_uniqueID))

all_trim$fourth_uniqueID<-paste(all_trim$cow_id, all_trim$quarter, "fourth", sep='_')#drop fourth_uniqueID so can have same column order as rest of , then add back in at end
```

```{r}
# error checking between old all_trim csv and new one- setdiff was helpful to tell differences in data sets. immogene vs. imogene for vontrapp was wrong, and fitzi13 (correct) vs fitzi31. somehow, these were not errors in original csv... maybe i had found them already? changed to imogene, and fitzi is 13 not 31

write.csv(all_trim, "all_trim_4_2_2021.csv", row.names = T)

test_alltrim<-read.csv("all_trim_test.csv", na.strings= ".", header = T)


vt_new<-filter(all_trim, first_farm== "vonTrapp" | second_farm == "vonTrapp" | third_farm== "vonTrapp")
vt_old<-filter(test_alltrim, first_farm== "vonTrapp" | second_farm == "vonTrapp" | third_farm== "vonTrapp")

pb_new<-filter(all_trim, first_farm== "Paddlebridge" | second_farm == "Paddlebridge" | third_farm== "Paddlebridge")
pb_old<-filter(test_alltrim, first_farm== "Paddlebridge" | second_farm == "Paddlebridge" | third_farm== "Paddlebridge")


new_cowid_quarter<-select(all_trim, c(cow_id, quarter))
old_cowid_quarter<-select(test_alltrim, c(cow_id, quarter))


setdiff(new_cowid_quarter, old_cowid_quarter)
setdiff(new_alltrim, all_trim1)

#old_cows<-(all_trim$cow_id)
new_cows<-(all_trim$cow_id)

setdiff(old_cows, new_cows)
setdiff(new_cows, old_cows)

# butter from OB had a space after her name for the first visit... so wasn't combining first visit with second and third
ob_new<-filter(all, first_farm== "OughtaBe" | second_farm == "OughtaBe" | third_farm== "OughtaBe")
```



#for some reason this alters colnames for OREI, so returning those to normal
```{r}
setwd(path.exercise)
OREI<-read.csv("OREIculture_results_csv.csv", na.strings= ".", header = T)
```

#do some checking to ensure farm visits subsetted correctly by checking cow IDs match
```{r}
table(OREI$visit)
#glimpse(first); glimpse(second); glimpse(third); glimpse(fourth)
table(first$first_farm); table(second$second_farm); table(third$third_farm); table(fourth$fourth_farm)

lf_all<-filter(all_trim, first_farm== "Lynd" | second_farm == "Lynd" | third_farm== "Lynd")
unique(lf_all$cow_id) #39 cows
lf_orei<- filter(OREI, farm=="Lynd")
unique(lf_orei$cow_id) #39 cows

pb_all<-filter(all_trim, first_farm== "Paddlebridge" | second_farm == "Paddlebridge" | third_farm== "Paddlebridge")
unique(pb_all$cow_id) #35 cows
pb_orei<- filter(OREI, farm=="Paddlebridge")
unique(pb_orei$cow_id) #35 cows

vt_all<-filter(all_trim, first_farm== "vonTrapp" | second_farm == "vonTrapp" | third_farm== "vonTrapp")
unique(vt_all$cow_id) #38 cows
vt_orei<- filter(OREI, farm=="vonTrapp")
unique(vt_orei$cow_id) #38 cows

gv_all<-filter(all_trim, first_farm== "Glennview" | second_farm == "Glennview" | third_farm== "Glennview")
unique(gv_all$cow_id) #37 cows
gv_orei<- filter(OREI, farm=="Glennview")
unique(gv_orei$cow_id) #37 cows

sw_all<-filter(all_trim, first_farm== "Swallowdale" | second_farm == "Swallowdale" | third_farm== "Swallowdale")
unique(sw_all$cow_id) #39 cows
sw_orei<- filter(OREI, farm=="Swallowdale")
unique(sw_orei$cow_id) #39 cows

sp_all<-filter(all_trim, first_farm== "StonyPond" | second_farm == "StonyPond" | third_farm== "StonyPond")
unique(sp_all$cow_id) #36 cows
sp_orei<- filter(OREI, farm=="StonyPond")
unique(sp_orei$cow_id) #36 cows

bj_all<-filter(all_trim, first_farm== "BJ" | second_farm == "BJ" | third_farm== "BJ")
unique(bj_all$cow_id) #39 cows
bj_orei<- filter(OREI, farm=="BJ")
unique(bj_orei$cow_id) #39 cows

bw_all<-filter(all_trim, first_farm== "Butterworks" | second_farm == "Butterworks" | third_farm== "Butterworks")
unique(bw_all$cow_id) #40 cows
bw_orei<- filter(OREI, farm=="Butterworks")
unique(bw_orei$cow_id) #40 cows

ob_all<-filter(all_trim, first_farm== "OughtaBe" | second_farm == "OughtaBe" | third_farm== "OughtaBe")
unique(ob_all$cow_id) #40 cows
ob_orei<- filter(OREI, farm=="OughtaBe")
unique(ob_orei$cow_id) #40 cows

cf_all<-filter(all_trim, first_farm== "Choiniere" | second_farm == "Choiniere" | third_farm== "Choiniere" | fourth_farm== "Choiniere")
unique(cf_all$cow_id) #41 cows
cf_orei<- filter(OREI, farm=="Choiniere")
unique(cf_orei$cow_id) #41 cows

```


##subset by a farm

```{r}
BW<-subset(OREI, farm =="Butterworks")
```

##how many isolates of everything do we have?

```{r}
library(dplyr)
count(OREI$path_1); count(OREI$path_2); count(OREI$path_3); count(OREI$path_4)
```

## chose most strict definitions of IMI yes in series for a farm
```{r}
bwIMI<-subset(BW, quarter_code == "1" | quarter_code == "2" | quarter_code == "21" | quarter_code == "24" | quarter_code == "27" | quarter_code == "29" | quarter_code =="35"); unique(bwIMI$quarter_code); count(bwIMI$quarter_code)
```

##check levels that all IMI series is either yes or no- should all be "yes" for ones that are subsetted as IMI

```{r}
unique(bwIMI$IMI_series);
which(bwIMI$IMI_series =="yes ")
```

##subset of ALL quarters with most strict definition of IMI
```{r}
OREI$quarter_code<-as.character(OREI$quarter_code)
strictIMI<-subset(OREI, quarter_code == "1" | quarter_code == "2" | quarter_code == "21" | quarter_code == "24" | quarter_code == "27" | quarter_code == "29" | quarter_code =="35")
unique(strictIMI$quarter_code) #checking right quarter codes selected
table(strictIMI$quarter_code) #seeing frequency of each type of quarter code
```


##check levels that all IMI series is either yes or no- should all be "yes" for ones that are subsetted as IMI

```{r}
unique(strictIMI$IMI_series)
table(strictIMI$IMI_series)
```


##how many isolates of everything do we have in STRICT IMI dataset?

```{r}
count(strictIMI$path_1); count(strictIMI$path_2); count(strictIMI$path_3); count(strictIMI$path_4)
```

## make spreadsheets for hygiene scoring stuff
```{r}
hygiene_OREI<-filter(OREI, quarter == "LH") #4212, down to 1053 (one line for each cow, vs 4 - multiple visits for same cow on seperate row)
hygiene_OREI<-select(hygiene_OREI, c(date, visit, farm, cow_id, hygiene_scorer_1, hygiene_1, hygiene_scorer_2, hygiene_2))

hygiene_all_trim<-filter(all_trim, quarter == "LH") #1536, down to 384 (one line for each cow, vs 4 - multiple visits for same cow on same row)
hygiene_all_trim<-select(hygiene_all_trim, c(cow_id, first_date, first_visit, first_farm, first_hygiene_scorer_1, first_hygiene_1, first_hygiene_scorer_2, first_hygiene_2, second_date, second_visit, second_farm, second_hygiene_scorer_1, second_hygiene_1, second_hygiene_scorer_2, second_hygiene_2, third_date, third_visit, third_farm, third_hygiene_scorer_1, third_hygiene_1, third_hygiene_scorer_2, third_hygiene_2, fourth_date, fourth_visit, fourth_farm, fourth_hygiene_scorer_1, fourth_hygiene_1, fourth_hygiene_scorer_2, fourth_hygiene_2))

write.csv(hygiene_OREI, "hygiene_from_OREI.csv", row.names = T)
write.csv(hygiene_all_trim, "hygiene_from_all_trim.csv", row.names = T)
```



